#Peer-graded Assignment:Course Project 1

##Loading and preprocessing of the data
```{r, echo = TRUE, results="hide"}
info<- read.csv("activity.csv")
s<- split(info,info$date)
```
## What is mean total number of steps taken per day
```{r}
nsteps<- lapply(s, function(x) sum(x$steps))
plotdata<- unlist(nsteps)
mean_of_steps <-round((mean(as.numeric(nsteps), na.rm = TRUE)), 0)
median_of_steps <- round((median(as.numeric(nsteps), na.rm = TRUE)),0)
hist(plotdata, xlab = "Number of Steps", main = "Histogram of the number of Steps taken per Day")
```




The mean of total number of steps taken by day iss `r mean_of_steps` and the median of the total number of steps taken by day is `r median_of_steps`


## What is the average daily activity pattern?
```{r,echo= TRUE}
times_series_data<- tapply(info$steps,info$interval,mean,na.rm= TRUE)
plot(row.names(times_series_data),times_series_data,type="l", xlab = "5 mins intervals", ylab = "Average steps", main = "Average Number of Steps Taken per Interval")
max<- which.max(times_series_data)
interval<- names(max)
```
The interval containing the averaged maximum number of steps is `r interval`


##Imputing Missing Values
```{r, echo=TRUE, results= " asis"}
na_counts <- sum(is.na(info$steps))
```
Number of rows with missing vaue is `r na_counts`

The strategy used to fill the missing value is using the mean for that 5-minute interval.

Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r, echo=TRUE, results= " hide"}
new_data <- info
for(i in 1:17568){
  if(is.na(info[i,1]))
new_data[i,]$steps<-times_series_data[as.character(info[i,]$interval)]
  
}
s_new <- split(new_data,new_data$date)
steps<- lapply(s_new, function(x) sum(x$steps))
data<- unlist(steps)
hist(data, xlab = "Number of steps", main = "Histogram of Total Number of Steps Taken per Day")
new_mean_of_steps <-round( mean(as.numeric(steps)),0)
new_median_of_steps <- round(median(as.numeric(steps)),0)
```
The mean value of the total steps taken per day with the new data set is `r new_mean_of_steps` and the medain of the total steps taken per day with the new data set is `r new_median_of_steps`.

The difference between the values from the data set with NA and the new dataset is not huge, little impact is observed.


## Are there differences in activity patterns between weekdays and weekends?
Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r, echo= TRUE}

new_data$day_type <- ""
days_week <- c("Montag","Dienstag","Mittwoch","Donnerstag","Freitag")

for(i in 1:17568){
  if(weekdays(as.Date(new_data[i,2])) %in% days_week){
    new_data[i,4] <- "weekday"
    
  }
  else{
     new_data[i,4] <- "weekend"
  }
  
}
new_data$day_type<- factor(new_data$day_type)
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r, echo=TRUE}
temp<- split(new_data,new_data$day_type)
plot_data1<- tapply((temp$weekday)$steps,(temp$weekday)$interval,mean,na.rm= TRUE)
plot_data2<-  tapply((temp$weekend)$steps,(temp$weekend)$interval,mean,na.rm= TRUE)

plot(row.names(times_series_data),plot_data1,type="l", xlab = "5 mins intervals", ylab = "Average steps in Weekdays", main = "Average Number of Steps Taken per Interval in weekdays")
plot(row.names(times_series_data),plot_data2,type="l", xlab = "5 mins intervals", ylab = "Average steps in Weekdend", main = "Average Number of Steps Taken per Interval in weekend")



```
